Gait Monitoring using Wireless Sensors\\ \vspace{4mm} \large{Final Report 18-748}\\ \large{Spring 2017}

### 2017-05-10 paper.tex
Added figures, more related work, more technical description

\usepackage{balance}
\usepackage{comment}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage{float}
\newcommand{\sys}{Safewalk\xspace}
\author{
  \IEEEauthorblockN{
   Iljoo Baek, Mengwen He and Emily Ruppel
  }
 \IEEEauthorblockA{
   Carnegie Mellon University
  }
 \IEEEauthorblockA{
   {ibaek,mengwenhe,eruppel}@andrew.cmu.edu
  }
}
\title{Gait Monitoring using Wireless Sensors\\
        \vspace{4mm}
        \large{Final Report 18-748}\\
        \large{Spring 2017}}

@@

\section{Technical Description}
The system architecture we employ to achieve our goal of measuring a subject's gait using
wireless sensors is shown in Figure~\ref{overview}. Note that it contains six Firefly
slave devices each with a Sparkfun Razor 9 DoF Razor IMU M0 board attached. The data are
then transmitted from the slave nodes to a master node which finally pushes the data to a
server. To implement our gait monitoring system we had to handle three major design
points- transmitting data from the IMUs, analyzing the IMU data, and building a useful
interface to the information. This section will describe the choices our team made with
regards to these three subsystems.

@@

\subsection{Gait Analysis}
  Paragraph about step detection

  Paragraph about stride length based on step detection and assuming the subject's limb
  measurements are known.

  Paragraph about calculating knee extension

@@

\section{Results}
Add picture of final enclosure. Discuss how it was attached to the subject. Mention that
it was fairly stable.

>>>>

\usepackage{balance}
\usepackage{comment}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage{float}
\usepackage{subfigure}
\newcommand{\sys}{Safewalk\xspace}
\author{
  \IEEEauthorblockN{
   Iljoo Baek, Mengwen He and Emily Ruppel
  }
 \IEEEauthorblockA{
   Carnegie Mellon University
  }
 \IEEEauthorblockA{
   {ibaek,mengwenhe,eruppel}@andrew.cmu.edu
  }
}
\title{Gait Monitoring using Wireless Sensors\\
        \vspace{4mm}
        \large{Final Report 18-748}\\
        \large{Spring 2017}}

@@

\subsection{IMU Data Processing}
A large amount of work has been performed in processing the acceleration and orientation
values reported by inertial measurement sensors to make use of the data. Specifically in
terms of gait analysis, the work comes from both the robotics and motion capture
communities. 
%Can we get a citation from the robotics community here too? 
For instance, Yost labs has characterized the different joint types of the
human body in terms of the necessary interpetation of data from IMUs mounted on adjacent
limbs~\cite{yost}. The complexity of the analysis ranges from the simple calculation to
determine the angle of a hinge joint to the vector transformations required to assess the
pitch, yaw and roll of multi-axis joints. Complementary work from the medical community
puts the raw measurements in context for clinicians by providing a point of comparison.
For instance, in ~\cite{pop} inertial sensor data was collected from a large population
of adults and translated into a database of typical gait parameter values that researchers
and healthcare professionals can use as a point of comparison. The study combines a
multitude of well characterized algorithms for assessing gait parameters using only two
IMUs, one on each of the subject's feet, and is able to build a model for normal gaits
in terms of parameters such as foot clearance and step cadence.

\section{Technical Description}
Our system follows in the spirit of prior work using IMUs to measure gait parameters, and
work that uses wireless sensors to improve the quality of patient care. The system
architecture we employ to achieve our goal of measuring a subject's gait using wireless
sensors is shown in Figure~\ref{overview}. Note that it contains six Firefly slave devices
each with a Sparkfun Razor 9 DoF Razor IMU M0 board attached. The data are then
transmitted from the slave nodes to a master node which finally pushes the data to a
server. To implement our gait monitoring system we had to handle three major design
points- transmitting data from the IMUs, analyzing the IMU data, and building a useful
interface to the information. This section will describe the choices our team made with
regards to these three subsystems.

@@

\subsection{Gait Analysis}
Our work analyzing subject's gait parameters focused on stride length, swing time, and
knee extension. Each of these parameters required different analysis techniques that will
be described below.\\ 
{\bf Stride Length} To determine the stride length, we first assume that the lengths of each
segment of a subject's limbs are known. Then...\\
%Need to know how this works!!
{\bf Swing Time} Our analysis focuses on detecting the subject's step to in turn ascertain
when a subject's foot is on the ground, and when it is in motion. By making the assumption
that the acceleration of the foot is close to zero when it is on the ground, we were able
to detect the beginning and end of a subject's stride based strictly on acceleration data
from the x-y plane. Our implementation is a simplified version of the technique used
in~\cite{fancy}. Figure~\ref{fig:step-detect} shows the basic shape of accelerometer data
gathered from a subject's feet as he or she walks and the marked regions indicate the time
during which his/her foot is either on the ground or in the air. 
\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/swing}
  \caption{{\bf Step detection from accelerometer data}}
  \label{fig:step-detect}
\end{figure}
\\
{\bf Knee Extension} A subject's knee extension was calculated using the hinge joint
techniques explained by Yost~\cite{yost}. The equations are: 
%Insert equations here!!

@@

\subsection{Enclosure} 
To ensure that the IMU remains in a fixed position on the body, and to protect the
connection between the Firefly and the IMU, we designed a 3D printed enclosure for the
slave nodes. Figure~\ref{fig:box} shows the enclosure with and without the specially designed
lid. The enclosure allows for easy access to all programming ports, power switches and
batteries without time consuming disassembly. Further, the design allows a strap to easily
pass through the upper and lower pieces of the enclosure so that the devices can be
comfortably affixed to subjects using velcro straps. During testing, we found that the
straps held the slave devices securely in place and they did not restrict the subject's
movements. 
%Except for that part where the ones on the thighs acted like tourniquets... but we won't
%talk about that... 
\begin{figure}[h]
  \centering
  \subfigure[]{
    \centering
    \includegraphics[width=.45\columnwidth]{figs/box}
    \label{fig:box2}
  }
  \subfigure[]{
    \centering
    \includegraphics[width=.45\columnwidth]{figs/wLid}
    \label{fig:wLid}
  }
  \label{fig:box}
  \caption{{\bf Enclosure for sensor nodes}}
\end{figure}

\section{Results}
